---
title: "P8105_HW1_yj2802"
author: "Yizhen Jia"
date: 2024-09-21
output: github_document
---


# Poblem 1

```{r}
data("penguins", package = "palmerpenguins")
```

Introduction:
The penguins dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
Including variables of `r names(penguins)`. It contains data from 2007 to 2009 of 3 species of penguins that are Adelie, Gentoo, and Chinstrap. 
The mean flipper length of the penguins is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`mm.


Below is the scatterplot of flipper_length_mm (y) vs bill_length_mm
(x):
```{r yx_scatter}
library(tidyverse)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +geom_point() +labs(title = "Flipper Length vs Bill Length", x = "Bill Length (mm)", y = "Flipper Length (mm)")
```

# Problem 2
Firstly, create 4 different variables:
```{r}
# set seed to ensure reproducibility
set.seed(1234)
# Create the data
numeric_var = rnorm(10)
logical_var = numeric_var > 0
character_var = letters[1:10]
factor_var = factor(rep(c("Level1", "Level2", "Level3"), length.out = 10))
# Combine into a dataframe
df = data.frame(numeric_var, logical_var, character_var, factor_var)
# Pull the variables column out of the dataframe
library(tidyverse)
numeric_vector = df %>% pull(numeric_var)
logical_vector = df %>% pull(logical_var)
character_vector = df %>% pull(character_var)
factor_vector = df %>% pull(factor_var)
```

(1) For a random sample of size 10 from a standard Normal distribution: Mean=`r mean(numeric_vector)`

(2) For a logical vector indicating whether elements of the sample are greater than 0: Mean=`r mean(logical_vector)`

(3) For a character vector of length 10: Mean=`r mean(character_vector)` 

(4) For a factor vector of length 10, with 3 different factor “levels”: Mean=`r mean(factor_vector)`

Result: the random sample and the logical vector works, while the character vector and the factor vector does not work.

When Convert the logical, character, and factor variables to numeric by the following code:

```{r}
numeric_logical <- as.numeric(df$logical_var)
numeric_character <- as.numeric(df$character_var)
numeric_factor <- as.numeric(df$factor_var)
```
(1) For a logical vector indicating whether elements of the sample are greater than 0: Mean=`r mean(numeric_logical)`. -->
This works because TRUE and FALSE are treated as 1 and 0, respectively.

(2) For a character vector of length 10: Mean=`r mean(numeric_character)`. -->
This gives an error because characters can’t be coerced into numbers.

(3) For a factor vector of length 10, with 3 different factor “levels”: Mean=`r mean(numeric_factor)`. -->
This works, but factors are converted to their underlying integer codes (e.g., Level1 --> 1, Level2 --> 2, etc.), so taking the mean might not make sense in this context.